---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.path = "man/figures/tidyexplain-",
  cache = TRUE
)
library(tidyexplain)
set_font_size(11, 26)
```

[gganimate]: https://github.com/thomasp85/gganimate#README
[dplyr-two-table]: https://dplyr.tidyverse.org/articles/two-table.html
[r4ds-set-ops]: http://r4ds.had.co.nz/relation-data.html#set-operations

# Tidy Animated Verbs

Garrick Aden-Buie -- [&commat;grrrck](https://twitter.com/grrrck) -- [garrickadenbuie.com](https://www.garrickadenbuie.com).
David Zimmermann -- [&commat;dav_zim](https://twitter.com/dav_zim) -- [datashenanigan.wordpress.com](https://datashenanigan.wordpress.com/)
Set operations contributed by [Tyler Grant Smith](https://github.com/TylerGrantSmith).

[![Binder](http://mybinder.org/badge.svg)](https://mybinder.org/v2/gh/gadenbuie/tidy-animated-verbs/master?urlpath=rstudio)
[![CC0](https://img.shields.io/badge/license_(images)_-CC0-green.svg)](https://creativecommons.org/publicdomain/zero/1.0/)
[![MIT](https://img.shields.io/badge/license_(code)_-MIT-green.svg)](https://opensource.org/licenses/MIT)

- Mutating Joins: [`inner_join()`](#inner-join), [`left_join()`](#left-join),
[`right_join()`](#right-join), [`full_join()`](#full-join)

- Filtering Joins: [`semi_join()`](#semi-join), [`anti_join()`](#anti-join)

- Set Operations: [`union()`](#union), [`union_all()`](#union-all), [`intersect()`](#intersect), [`setdiff()`](#setdiff)

- Tidyr Operations: [`gather()`](#gather), [`spread()`](#spread)

- Learn more about
    - [Relational Data](#relational-data)
    - [gganimate](#gganimate)


Please feel free to use these images for teaching or learning about action verbs from the [tidyverse](https://tidyverse.org).
You can directly download the [original animations](images/) or static images in [svg](images/static/svg/) or [png](images/static/png/) formats, or you can use the [scripts](R/) to recreate the images locally.

Currently, the animations cover the [dplyr two-table verbs][dplyr-two-table] and I'd like to expand the animations to include more verbs from the tidyverse.
[Suggestions are welcome!](https://github.com/gadenbuie/tidy-animated-verbs/issues)

## Installing

The in-development version of `tidyexplain` can be installed with `devtools`:

```r
# install.package("devtools")
devtools::install_github("gadenbuie/tidy-animated-verbs")

library(tidyexplain)
```

## Mutating Joins

```{r intial-dfs}
x <- dplyr::data_frame(
  id = 1:3,
  x = paste0("x", 1:3)
)

y <- dplyr::data_frame(
  id = (1:4)[-3],
  y = paste0("y", (1:4)[-3])
)

animate_full_join(x, y, by = c("id"), export = "first")
```


```{r}
x
y
```

### Inner Join

> All rows from `x` where there are matching values in `y`, and all columns from `x` and `y`.

```{r inner-join}
animate_inner_join(x, y, by = "id")
```


```{r}
dplyr::inner_join(x, y, by = "id")
```

### Left Join

> All rows from `x`, and all columns from `x` and `y`. Rows in `x` with no match in `y` will have `NA` values in the new columns.

```{r left-join}
animate_left_join(x, y, by = "id")
```


```{r}
dplyr::left_join(x, y, by = "id")
```

### Left Join (Extra Rows in y)

> ... If there are multiple matches between `x` and `y`, all combinations of the matches are returned.

```{r left-join-extra}
y_extra <- dplyr::bind_rows(y, dplyr::data_frame(id = 2, y = "y5"))
y_extra # has multiple rows with the key from `x`

animate_left_join(x, y_extra, by = "id", 
  anim_opts = anim_options(title_size = 22))
```

```{r}
dplyr::left_join(x, y_extra, by = "id")
```

### Right Join

> All rows from y, and all columns from `x` and `y`. Rows in `y` with no match in `x` will have `NA` values in the new columns.

```{r right-join}
animate_right_join(x, y, by = "id")
```


```{r}
dplyr::right_join(x, y, by = "id")
```

### Full Join

> All rows and all columns from both `x` and `y`. Where there are not matching values, returns `NA` for the one missing.

```{r full-join}
animate_full_join(x, y, by = "id")
```


```{r}
dplyr::full_join(x, y, by = "id")
```

## Filtering Joins

### Semi Join

> All rows from `x` where there are matching values in `y`, keeping just columns from `x`.

```{r semi-join}
animate_semi_join(x, y, by = "id")
```


```{r}
dplyr::semi_join(x, y, by = "id")
```

### Anti Join

> All rows from `x` where there are not matching values in `y`, keeping just columns from `x`.

```{r anti-join}
animate_anti_join(x, y, by = "id")
```


```{r}
dplyr::anti_join(x, y, by = "id")
```

## Set Operations

```{r intial-dfs-so}
x <- dplyr::data_frame(
  x = c(1, 1, 2),
  y = c("a", "b", "a")
)
y <- dplyr::data_frame(
  x = c(1, 2),
  y = c("a", "b")
)

animate_union(x, y, export = "first")
```


```{r}
x
y 
```

### Union

> All unique rows from `x` and `y`.

```{r union}
animate_union(x, y)
```


```{r}
dplyr::union(x, y)
```



```{r union-y-x}
animate_union(y, x)

dplyr::union(y, x)
```

### Union All

> All rows from `x` and `y`, keeping duplicates.

```{r union-all}
animate_union_all(x, y)
```



```{r}
dplyr::union_all(x, y)
```


### Intersection

> Common rows in both `x` and `y`, keeping just unique rows.

```{r intersect}
animate_intersect(x, y)
```


```{r}
dplyr::intersect(x, y)
```

### Set Difference

> All rows from `x` which are not also rows in `y`, keeping just unique rows.

```{r setdiff}
animate_setdiff(x, y)
```


```{r}
dplyr::setdiff(x, y)
```


```{r setdiff-y-x}
animate_setdiff(y, x)

dplyr::setdiff(y, x)
```

## Tidy Data and `gather()`, `spread()` functionality

[Tidy data](http://r4ds.had.co.nz/tidy-data.html#tidy-data-1) follows
the following three rules:

1.  Each variable has its own column.
2.  Each observation has its own row.
3.  Each value has its own cell.

Many of the tools in the [tidyverse](https://tidyverse.org) expect data
to be formatted as a tidy dataset and the
[tidyr](https://tidyr.tidyverse.org) package provides functions to help
you organize your data into tidy data.

```{r}
long <- dplyr::data_frame(
  year = c(2010, 2011, 2010, 2011, 2010, 2011),
  person = c("Alice", "Alice", "Bob", "Bob", "Charlie", "Charlie"),
  sales = c(105, 110, 100, 97, 90, 95)
)
wide <- dplyr::data_frame(
  year = 2010:2011, 
  Alice = c(105, 110), 
  Bob = c(100, 97), 
  Charlie = c(90, 95)
)
```

### Gather

> Gather takes multiple columns and collapses into key-value pairs, duplicating all other columns as needed. You use gather() when you notice that your column names are not names of variables, but values of a variable.

```{r gather}
set_font_size(4.5, 15)
animate_gather(wide, key = "person", value = "sales", -year)
```

```{r}
tidyr::gather(wide, key = "person", value = "sales", -year)
```

### Spread

> Spread a key-value pair across multiple columns. Use it when an a column contains observations from multiple variables.

```{r spread}
animate_spread(long, key = "person", value = "sales")
```

```{r}
tidyr::spread(long, key = "person", value = "sales")
```


## Learn More

### Relational Data

The [Relational Data](http://r4ds.had.co.nz/relation-data.html) chapter of the
[R for Data Science](http://r4ds.had.co.nz/) book by Garrett Grolemund and Hadley Wickham
is an excellent resource for learning more about relational data.

The [dplyr two-table verbs vignette][dplyr-two-table]
and Jenny Bryan's [Cheatsheet for dplyr join functions](http://stat545.com/bit001_dplyr-cheatsheet.html)
are also great resources.

### gganimate

The animations were made possible by the newly re-written [gganimate] package by 
[Thomas Lin Pedersen](https://github.com/thomasp85)
(original by [Dave Robinson](https://github.com/dgrtwo)).
The [package readme][gganimate] provides an excellent (and quick) introduction to gganimte.
